Automated Formal Synthesis of Digital Controllers for State-Space Physical Plants

### 2017-01-23 cav2017/cav2017.tex
more revision of the AA based approach

Let us recall the formulas used for single-input and single-output (SISO) systems 
in sections \ref{sec:reachable} and \ref{sec:observable}.
A controllable system with quantization will have the following dynamics:
\todo{eliminate vectors}
%
\begin{align}
\label{eq:observer_LTI_cf}
\vec{x}_{k+1}=&\mat{A}_{t}\vec{x}_k+\mat{B}_{tp} \vec{r}_k+\mat{B}_{tn} (\vec{\nu}_1 + \vec{\nu}_2 + \vec{\nu}_3)
\end{align}
\begin{align*}
\mat{A}_{t}=&\mat{T}\mat{A}_d\mat{T}^{-1}&
\mat{B}_{tp}=&\mat{T}\mat{B}_d&
\mat{B}_{tn}=&\mat{T}\mat{B}_d&
\mat{T}=&\mat{R}_{cf}\mat{R}^{-1}&
\end{align*}

\begin{align*}
\mat{A}_{t}=\mat{T}\mat{A}_d\mat{T}^{-1}=&\left[
\begin{array}{ccccc}
0&1&0&\cdots&0\\
0&0&1&\cdots&0\\
\vdots&\vdots&\vdots&\ddots&\vdots\\
0&0&0&\cdots&1\\
k_n-a_n&k_{n-1}-a_{n-1}&k_{n-2}-a_{n-2}&\cdots&k_1-a_1
\end{array}\right]
\label{eq:cf_SISO_2}
\end{align*}

The characteristic polynomial of the above matrix is:
%
\begin{align*}
z^n+\sum_{i=1}^p{(a_i-k_i)z^{n-i}}
\end{align*}

For a stabilized closed-loop, which is necessary for our safety property, we need
these roots to remain within the unit circle, which we verify using Jury's criteria.
Notice that in our polynomials $c_0=1$ and we only use reals.

Once we have established convergence, we may examine the 
specification of the state and input spaces ($\phi_{safety} \wedge \phi_{input}$).
Whilst there are other properties, we may include in the verification, 
we will solely focus on these for simplicity of the explanation.
We assume that we do not want saturation in our system, 
hence the control loop has only one mode. We have also identified 
the quantization noise as a bounded non-deterministic input.
%
The first step constraints
\begin{equation}
\mat{K} \models \vec{u}_k=-\mat{K}\vec{x}_k : \phi_{safety} \wedge \phi_{input}
\label{eq:of_bounds}
\end{equation}
The second step is evaluating the progression of the system over time to evaluate 
$\vec{x}_{k+1} \models \phi_{safety}$. For this step we require 
an initial set from which the system progresses. Obviously, this set $\vec{x}_0$ 
must satisfy the specification; otherwise, the system will be unsafe to begin with.
We accept a specification in the form $\mat{E}\mat{T}\vec{x}_0<\mat{f}$. 
The presence of $\mat{T}$ is because this will typically be given in the original 
state-space. Accelerating \eqref{eq:observer_LTI_cf}, we obtain:
%
\begin{align}
\label{eq:acc_observer_LTI_cf}
\vec{x}=&\mat{A}_t^k\vec{x}_0
+\sum_{i=0}^{k-1} \mat{A}_t^i \mat{B}_{tp} \vec{r}_i
+\sum_{i=0}^{k-1} \mat{A}_t^i \mat{B}_{tn}(\vec{\nu}_1+\vec{\nu}_2+\vec{\nu}_3) : \vec{x}_0 \models \phi_{init}
\end{align}
which would still require us to verify the system for every $k$ up to infinity. 
Instead we use abstract acceleration~\cite{cattaruzza2015unbounded}, which translates the formula into
%
\begin{align}
\label{eq:aa_observer_LTI_cf}
\hat{X}^\#
=&\mathcal{A}_t X_0+\mathcal{B}_{tp} R + \mathcal{B}_{tn} N\\
X_0 =&\left \{\vec{x} : \vec{x} \models \phi_{init} \right\}\nonumber\\
 R =&\{r : \mat{E}_r < \vec{f}_r \}\nonumber\\
 N=&\left \{ \vec{\nu}_1+\vec{\nu}_2+ \vec{\nu}_3 : \nu_1 \in \left[-\frac{q1}{2}\ \ \frac{q1}{2}\right] \wedge \nu_2 \in \left[-\frac{q2}{2}\ \ \frac{q2}{2}\right]  \wedge \nu_3 \in \left[-q3\ \ q3\right]  \right \}\nonumber
\end{align}
%
where $\mathcal{A}_t=\bigcup_{k=1}^\infty \mat{A}^k, \mathcal{B}_{tp}=\bigcup_{k=1}^\infty \sum_{i=0}^k\mat{A}^i\mat{B}_{tp}, \mathcal{B}_{tn}=\bigcup_{k=1}^\infty \sum_{i=0}^k\mat{A}^i\mat{B}_{tn}$ are abstract matrices for the system in \eqref{eq:observer_LTI_cf} and $r$ is a non-deterministic parametric input (\emph{ie} $\forall k>0, r_{k+1} =r_k$) whereas the set N is composed of non-deterministic variable inputs.

>>>>

% Let us recall the formulas used for single-input and single-output (SISO) systems 
% in sections \ref{sec:reachable} and \ref{sec:observable}.
% A controllable system with quantization will have the following dynamics:
% \todo{eliminate vectors}
% %
% \begin{align}
% \label{eq:observer_LTI_cf}
% \vec{x}_{k+1}=&\mat{A}_{t}\vec{x}_k+\mat{B}_{tp} \vec{r}_k+\mat{B}_{tn} (\vec{\nu}_1 + \vec{\nu}_2 + \vec{\nu}_3)
% \end{align}
% \begin{align*}
% \mat{A}_{t}=&\mat{T}\mat{A}_d\mat{T}^{-1}&
% \mat{B}_{tp}=&\mat{T}\mat{B}_d&
% \mat{B}_{tn}=&\mat{T}\mat{B}_d&
% \mat{T}=&\mat{R}_{cf}\mat{R}^{-1}&
% \end{align*}

% \begin{align*}
% \mat{A}_{t}=\mat{T}\mat{A}_d\mat{T}^{-1}=&\left[
% \begin{array}{ccccc}
% 0&1&0&\cdots&0\\
% 0&0&1&\cdots&0\\
% \vdots&\vdots&\vdots&\ddots&\vdots\\
% 0&0&0&\cdots&1\\
% k_n-a_n&k_{n-1}-a_{n-1}&k_{n-2}-a_{n-2}&\cdots&k_1-a_1
% \end{array}\right]
% \label{eq:cf_SISO_2}
% \end{align*}

% The characteristic polynomial of the above matrix is:
% %
% \begin{align*}
% z^n+\sum_{i=1}^p{(a_i-k_i)z^{n-i}}
% \end{align*}

% For a stabilized closed-loop, which is necessary for our safety property, we need
% these roots to remain within the unit circle, which we verify using Jury's criteria.
% Notice that in our polynomials $c_0=1$ and we only use reals.
\todo{where is $\phi_{input}$ defined?}
Once we have established convergence, we may examine the 
specification of the state and input spaces ($\phi_{safety} \wedge \phi_{input}$).
% Whilst there are other properties, we may include in the verification, 
% we will solely focus on these for simplicity of the explanation.
% We assume that we do not want saturation in our system, 
% hence the control loop has only one mode.

We have also identified 
the quantization noise as a bounded non-deterministic input.
%
The first step constraints
\begin{equation}
\mat{K} \models \vec{u}_k=-\mat{K}\vec{x}_k : \phi_{safety} \wedge \phi_{input}
\label{eq:of_bounds}
\end{equation}
The second step is evaluating the progression of the system over time to evaluate 
$\vec{x}_{k+1} \models \phi_{safety}$. For this step we require 
an initial set from which the system progresses. Obviously, this set $\vec{x}_0$ 
must satisfy the specification; otherwise, the system will be unsafe to begin with.
We accept a specification in the form $\mat{E}\mat{T}\vec{x}_0<\mat{f}$. 
The presence of $\mat{T}$ is because this will typically be given in the original 
state-space. Accelerating \eqref{eq:observer_LTI_cf}, we obtain:
%
\begin{align}
\label{eq:acc_observer_LTI_cf}
\vec{x}=&\mat{A}_t^k\vec{x}_0
+\sum_{i=0}^{k-1} \mat{A}_t^i \mat{B}_{tp} \vec{r}_i
+\sum_{i=0}^{k-1} \mat{A}_t^i \mat{B}_{tn}(\vec{\nu}_1+\vec{\nu}_2+\vec{\nu}_3) : \vec{x}_0 \models \phi_{init}
\end{align}
which would still require us to verify the system for every $k$ up to infinity. 
Instead we use abstract acceleration~\cite{cattaruzza2015unbounded}, which translates the formula into
%
\begin{align}
\label{eq:aa_observer_LTI_cf}
\hat{X}^\#
=&\mathcal{A}_t X_0+\mathcal{B}_{tp} R + \mathcal{B}_{tn} N\\
X_0 =&\left \{\vec{x} : \vec{x} \models \phi_{init} \right\}\nonumber\\
 R =&\{r : \mat{E}_r < \vec{f}_r \}\nonumber\\
 N=&\left \{ \vec{\nu}_1+\vec{\nu}_2+ \vec{\nu}_3 : \nu_1 \in \left[-\frac{q1}{2}\ \ \frac{q1}{2}\right] \wedge \nu_2 \in \left[-\frac{q2}{2}\ \ \frac{q2}{2}\right]  \wedge \nu_3 \in \left[-q3\ \ q3\right]  \right \}\nonumber
\end{align}
%
where $\mathcal{A}_t=\bigcup_{k=1}^\infty \mat{A}^k, \mathcal{B}_{tp}=\bigcup_{k=1}^\infty \sum_{i=0}^k\mat{A}^i\mat{B}_{tp}, \mathcal{B}_{tn}=\bigcup_{k=1}^\infty \sum_{i=0}^k\mat{A}^i\mat{B}_{tn}$ are abstract matrices for the system in \eqref{eq:observer_LTI_cf} and $r$ is a non-deterministic parametric input (\emph{ie} $\forall k>0, r_{k+1} =r_k$) whereas the set N is composed of non-deterministic variable inputs.
